---
title: "Lecture 7: Live Code Day: Attacking a Messy Spreadsheet!"
author: "Wenna Xi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(table1)
```

## Announcements 

* Data Challenge #2 is due Monday 10/9 @ 5:59pm on Canvas
* Quiz #3 is on Monday 10/29 at 6pm
* UCSAS 2024 USOPC DATA CHALLENGE: https://statds.org/events/ucsas2024/challenge.html
* Github repo

## Live Code Tasks

The `ehr.xlsx` file contains simulated EHR data I generated based on a real project

The file has three tabs: 

* demographics: patients' admission date/time, discharge date/time, visit type (Inpatient, ED), hospital sites (LMH, Queens)
* medications: date/time, dose of sedating psychotropic medications administered during the visit
* observation: start and end date/time when patient was placed under one-on-one safety observation 

Tasks:

1. For each drug $\times$ route combination, calculate the total dose administered for each patient during each visit
2. Unique number of agitation episodes requiring pharmacologic intervention
    - Use number of discrete administrations of sedating psychotropic medications as an approximation 
    - Discrete administration defined as two administrations > 10 minutes apart 
3. Total length of time patients were placed under observation during each visit




## Live Code

### Read in the data

```{r}
demographics <- read_excel(here("data", "ehr.xlsx"), sheet = "demographics") |>
  clean_names()
medications <- read_excel(here("data", "ehr.xlsx"), sheet = "medications") |>
  clean_names()
observation <- read_excel(here("data", "ehr.xlsx"), sheet = "observation") |>
  clean_names()
```

### Task 1
```{r}
table(medications$drug_name, useNA = "always")
table(medications$task_route, useNA = "always")
table(medications$dose_unit, useNA = "always")


medications_task1 <- medications |>
  mutate(drug_new = word(str_to_sentence(drug_name)),
         route_new = if_else(task_route == "IntraMUSCULAR", "IM", "IV"),
         drug_route = paste(drug_new, route_new, sep = "_")) |>
  group_by(patient_id, visit_id, drug_route) |>
  summarise(total_dose = sum(dose)) |>
  pivot_wider(names_from = drug_route,
              values_from = total_dose)
```

### Task 2
```{r}
medications_task2 <- medications |>
  group_by(patient_id, visit_id) |>
  arrange(date) |>
  mutate(next_date = lead(date),
         diff = interval(date, next_date) / minutes(1),
         unique_episodes = if_else(diff > 10, 1, 0)) |>
  filter(unique_episodes == 1 | is.na(unique_episodes)) |>
  summarise(n_unique_espisodes = n())

```


### Task 3
```{r}
observation_task3 <- observation |>
  group_by(patient_id, visit_id) |>
  arrange(start_time, end_time) |>
  mutate(next_start_time = lead(start_time),
         nonoverlap__observations = if_else(end_time > next_start_time, 1, 0),  
         distinct_observations = c(0, cumsum(as.numeric(next_start_time) > cummax(as.numeric(end_time)))[-n()])) |>
  group_by(patient_id, visit_id, distinct_observations) |>
  summarise(start = min(start_time), 
            end = max(end_time),
            length = interval(start, end) / hours(1))
```

### Table Summary
```{r}
results <- demographics |>
  left_join(medications_task1) |>
  left_join(medications_task2) |>
  left_join(observation_task3)

table1(~ Diphenhydramine_IM + Diphenhydramine_IV + Haloperidol_IV + Lorazepam_IV + n_unique_espisodes + length | hospital, data = results)
```
